EuPPollNet: A European database of plant-pollinator networks

### 2024-05-09 Manuscript/EuPPollNet_Ms.qmd
edits

Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe. These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses. In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions. Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations. EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods. In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time. We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites. The solid black points indicate the number of species and interactions contained in the database. Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel. Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites. Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}

@@

#Check number of studies by habitat
habitat = data %>% 
select(Study_id, EuPPollNet_habitat) %>% 
distinct() %>%  
group_by(EuPPollNet_habitat) %>% 
summarise(Habitat_count = n()) 
#Check number of studies by bioregion
bioregion = data %>% 
select(Study_id, Bioregion) %>% 
distinct() %>%  
group_by(Bioregion) %>% 
summarise(Bioregion_count = n()) 
#Check pols by habitat
pollinator_percent = data %>%
group_by(EuPPollNet_habitat, Pollinator_order) %>%
filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
filter(Pollinator_order %in% main_orders) %>% 
summarise(count = n()) %>%
group_by(EuPPollNet_habitat) %>%
mutate(Proportion = count/sum(count)) %>%
select(-count) 
#Check pols by bioregion
pollinator_percent_bio = data %>%
group_by(Bioregion, Pollinator_order) %>%
filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
filter(Pollinator_order %in% main_orders) %>% 
summarise(count = n()) %>%
group_by(Bioregion) %>%
mutate(Proportion = count/sum(count)) %>%
select(-count) 
#Preparing datasets for horizontal plot
a1 = habitat %>% 
mutate(Group = "Habitat") %>% 
rename(Habitat = EuPPollNet_habitat) %>% 
rename(Count = Habitat_count)

@@

Connectance and nestedness did not change across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**). However, differences were found for connectance but not nestedness across bioclimatic regions. Specifically, the alpine region differed statistically in connectance to the other bioclimatic regions (Tukey test _P_ < 0.01). In addition, the atlantic and continental bioclimatic regions also showed statistical meaningful differences in connectance (Tukey test _P_ < 0.01). Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**). Finally, we found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6d**).

@@

Things to do:
-Habitat Figure: Patterns of boreal are weird --> Comment in results or discussion
-Figure 2 Bombus por encima de Apis, explain that Apis is not included
-Intro: Somewhere, bur maybe later, we should mention what we add. 1) Way more networks than those on Mangal or Webof life (which you should cite). 2) curated metadata and harmonization, 3) Community level data and not only isolated interactions as in GloBI.
-Network properties / significance issue

>>>>

Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe. These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses. In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions. Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations. EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods. In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time. This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions. We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites. The solid black points indicate the number of species and interactions contained in the database. Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel. Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites. Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site. Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}

@@

#Karise is a very large study that only sampled bumblebees and
#is providing unrealistic results for the Boreal region
#Filter out
#Maybe add to caption?
data1 = data %>% 
filter(!Study_id == "13_Karise")

#Check number of studies by habitat
habitat = data1 %>% 
select(Study_id, EuPPollNet_habitat) %>% 
distinct() %>%  
group_by(EuPPollNet_habitat) %>% 
summarise(Habitat_count = n()) 
#Check number of studies by bioregion
bioregion = data1 %>% 
select(Study_id, Bioregion) %>% 
distinct() %>%  
group_by(Bioregion) %>% 
summarise(Bioregion_count = n()) 
#Check pols by habitat
pollinator_percent = data1 %>%
group_by(EuPPollNet_habitat, Pollinator_order) %>%
filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
filter(Pollinator_order %in% main_orders) %>% 
summarise(count = n()) %>%
group_by(EuPPollNet_habitat) %>%
mutate(Proportion = count/sum(count)) %>%
select(-count) 
#Check pols by bioregion
pollinator_percent_bio = data1 %>%
group_by(Bioregion, Pollinator_order) %>%
filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
filter(Pollinator_order %in% main_orders) %>% 
summarise(count = n()) %>%
group_by(Bioregion) %>%
mutate(Proportion = count/sum(count)) %>%
select(-count) 
#Preparing datasets for horizontal plot
a1 = habitat %>% 
mutate(Group = "Habitat") %>% 
rename(Habitat = EuPPollNet_habitat) %>% 
rename(Count = Habitat_count)

@@

Connectance and nestedness did not show relevant trends across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**). However, differences were found for connectance but not nestedness across bioclimatic regions. Specifically, the alpine region differed statistically in connectance to the other bioclimatic regions (Tukey test _P_ < 0.01). In addition, the atlantic and continental bioclimatic regions also showed statistical meaningful differences in connectance (Tukey test _P_ < 0.01). Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**). Finally, we found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6d**).

@@
